---
title: "Blog Post 6: Thesis and Continuing Exploration"
author: "Group 10"
date: "2024-04-19"
date-modified: "2024-04-19"
draft: FALSE
---
To narrow down to certain states, we will focus on Massachusetts and Mississippi. Massachusetts is the most educated state in the US, and Mississippi is the least educated state in 2009, when our data starts. 

## Thesis
Mississippi, with its higher percentage of economically disadvantaged students, higher poverty rate, higher unemployment rate, lower percentage of parents holding bachelor's degrees, and lower median income, has lower mean scores in both language and math for all students compared to Massachusetts.

## More Exploration of Independent Variables
### Massachusetts
```{r, echo=FALSE}
library(tidyverse)
library(dplyr)
library(readr)
cleaned_seda <- read_rds("dataset/cleaned_seda.rds")
massachusetts_data <- cleaned_seda %>%
  filter(state == "MA")
mississippi_data <- cleaned_seda %>%
  filter(state == "MS")

selected_variables_MA <- massachusetts_data %>%
  select(mn_score_all, perecd, povertyall, unempall, lninc50all, baplusall, perblk, perasn, perwht) %>%
  filter(if_all(everything(), is.finite))

selected_variables_MA %>%
  gather(key = "variable", value = "score") %>%
  ggplot(aes(x = score)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black") +
  facet_wrap(~ variable, scales = "free") +
  labs(title = "Distribution of Variables for MA Districts",
       x = "Test Score",
       y = "Frequency") +
  theme_minimal()
```

### Mississippi
```{r, echo=FALSE}
selected_variables_MS <- mississippi_data %>%
  select(mn_score_all, perecd, povertyall, unempall, lninc50all, baplusall, perblk, perasn, perwht) %>%
  filter(if_all(everything(), is.finite))

selected_variables_MS %>%
  gather(key = "variable", value = "score") %>%
  ggplot(aes(x = score)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black") +
  facet_wrap(~ variable, scales = "free") +
  labs(title = "Distribution of Variables for MS Districts",
       x = "Test Score",
       y = "Frequency") +
  theme_minimal()
```
After observing each plot for the variables, all variables are skewed except for the median income and the mean score. Therefore, we will apply natural log or square root transformations for those variables to deal with the skewness.

```{r, echo=FALSE}
library(tidyverse)

selected_variables_MS <- mississippi_data %>%
  select(perasn, perblk, perecd, perwht, unempall, baplusall) %>%
  filter(perasn > 0 & perblk > 0 & perecd > 0 & perwht > 0 & unempall > 0)

selected_variables_MS_log <- selected_variables_MS %>%
  mutate(across(everything(), log))

selected_variables_MS_sqrt <- selected_variables_MS %>%
  mutate(across(everything(), sqrt))

selected_variables_MS_log %>%
  gather(key = "variable", value = "value") %>%
  ggplot(aes(x = value)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black") +
  facet_wrap(~ variable, scales = "free") +
  labs(title = "Histograms of Log-Transformed Variables for MS Districts",
       x = "Log-transformed Value",
       y = "Frequency") +
  theme_minimal()

selected_variables_MS_sqrt %>%
  gather(key = "variable", value = "value") %>%
  ggplot(aes(x = value)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black") +
  facet_wrap(~ variable, scales = "free") +
  labs(title = "Histograms of Square Root Transformed Variables for MS Districts",
       x = "Square Root Transformed Value",
       y = "Frequency") +
  theme_minimal()
```

Since some of the variables are still skewed, we will find more potential transformations or use a Robust Regression model that is less sensitive to outliers and skewness.