---
title: "Blog post 4: Exploring the Data"
author: "Group 10"
date: "2024-04-08"
date-modified: "2024-04-08"
draft: FALSE
---

## Our Trends

As seen in blog post 3, race was a predictor variable in the scores seen in students. Economic status is another. We want to see how these two variables work together, rather than individually like we did last week. 

----

## More Exploration



----

## simple linear regression model (mean score of all students ~ percentage of economically disadvantaged student )

```{r, echo=FALSE}
library(dplyr)
library(ggplot2)
library(broom)
library(tidyverse)

ds = cleaned_seda

lm_model <- lm(mn_score_all ~ perecd, data = ds)
summary(lm_model)
```
Coefficient (Beta1) = -1.208213: For every one-unit increase in the percentage of economically disadvantaged students, the mean score of all students decreases by approximately 1.208213 units. R-squared = 0.5674: About 56.74% of the variability in student mean scores is explained by the percentage of economically disadvantaged students. P-value for predictor < 2e-16: The negative relationship between the percentage of economically disadvantaged students and the mean score of all students is statistically significant, with a p-value less than 2e-16.

----

## Residuals and Related Statistics Table:
```{r echo=FALSE, message=FALSE, warning=FALSE}
augment_lm <- augment(lm_model)
kable(head(augment_lm), format = "pipe")
```

----

## Residual Plot
```{r}
# Fit the model
residuals <- resid(mlr_model)
fitted_values <- fitted(mlr_model)


plot(fitted_values, residuals,
     xlab = "Fitted Values",
     ylab = "Residuals",
     main = "Residual Plot of mlr_model",
     pch = 20)

abline(h = 0, col = "red")
```

